Blum axioms and nondeterministic computation of functions

### 2016-03-21 paper.tex
paper.tex: Rewrite introduction.

\address{
    Departamento de Informática e Estatística
    --- Universidade Federal de Santa Catarina
    \email{royertiago@gmail.com}
}

@@

\section{Gödel numberings}

@@

\section{Nondeterministic computation of functions}

>>>>

\address{
    UFSC/USP?
}

@@

In Theory of Computation,
we usually use languages
to mathematically model problems in the real world.
Decision problems (``yes/no'') are mapped to languages in a very natural way:
just put every ``yes'' instance in the language,
and leave the rest out.
Search problems usually are rewritten as a decision problem,
and then this problem is converted to a language.
For instance,
the task of finding a satisfying assignment for a Boolean formula $\phi$
is reinterpreted to the task of merely deciding whether this assignment exists,
and this task is then converted to a language
--- in this example we have $\SAT$, the Boolean satisfiability problem.
This does the trick when it comes to proving that something is hard;
for instance, if we show that the decision problem is $\NP$-hard or undecidable,
then intuitively the corresponding search problem must be at least as hard.
Therefore,
concepts like ``decidable'', ``$\NP$-complete'', ``polynomial-time computable''
only arise in the context of decision problems.

However, even from the theoretical standpoint,
it is useful to extend these concepts to functions.
For instance,
the concept of polynomial-time computable functions
are central to the definition of Karp reductions.
For single-tape, deterministic Turing machines,
this definition is easy to extend:
A function $f : \{0, 1\}^* \to \{0, 1\}^*$
is said to be \emph{polynomial-time computable}
if there is a Turing machine $M$ and a polynomial $p$ such that,
when given the string $x$ as input,
$M$ halts with $f(x)$ in its tape within $p(|x|)$ steps.

Most extensions to this basic model,
like the use of several tracks, or several tapes, or multidimensional tapes,
can be easily incorporated in the definition of polynomial-time computable.
Except nondeterminism.%
\footnote{
    Note also all the other extensions can be simulated by the original model
    suffering at most a polynomial slowdown,
    even in the context of decision problems.
    This is seems not to be the case with nondeterminism.
}
We hit a wall right in the start:
how does a nondeterministic machine computes a function in the first place?

This paper presents several attempts to establish
how a nondeterministic Turing machine can compute a function,
and to extend the concept of ``polynomial-time computable'' under each definition.
Sections \ref{sec:godel_numberings} and~\ref{sec:blum-axioms} sets reasonableness criteria
to both the definition of function computation and complexity of the computation.
Section~\ref{sec:nondeterministic_computation_of_functions}
presents and criticizes the several definitions.

\section{Gödel numberings}
\label{sec:godel_numberings}

@@

\section{Nondeterministic computation of functions}
\label{sec:nondeterministic_computation_of_functions}
